---
author: "Simon Garnier"
title: "1 - Installing ROpenCVLite"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{1 - Installing ROpenCVLite}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Installing [`ROpenCVLite`](https://github.com/swarm-lab/ROpenCVLite) should go 
smoothly on all major operating systems (Mac, Windows, and most major Linux 
distributions). There are a few steps to follow before installing the package 
itself in order to prepare all the necessary external (i.e. non-R) dependencies 
for [`OpenCV`](https://opencv.org/). 

All installation steps are described below. 

---

+ [1 - Before installing ](#before-installing )
  + [1.1 - Windows](#windows)
  + [1.2 - Mac](#mac)
  + [1.3 - Linux](#linux)
  + [1.4 - Mac, Linux and Windows](#mac-linux-and-windows)
+ [2 - Installing `ROpenCVLite`](#installing-ropencvlite)
  
---

## 1 - Before installing 

### 1.1 - Windows

Before installing `ROpenCVLite`, you will need to install the latest versions of
[`CMake`](https://cmake.org/) and [`Rtools`](https://cran.r-project.org/bin/windows/Rtools/). 

#### 1.1.1 - Cmake

+ **Option 1**

Use the [`winget`](https://learn.microsoft.com/en-us/windows/package-manager/winget/) 
utility to install `Cmake` from the terminal/PowerShell:

```{powershell eval=FALSE}
winget install cmake
```

+ **Option 2**

Use the convenience function provided by the `installr` library: 

```{r eval=FALSE}
if (!require("installr"))
  install.packages("installr")
  
installr::install.CMake()
```

During the installation process, make sure to tell the installer to add `CMake` 
to your "PATH".

+ **Option 3**

Download the `CMake` installer for Windows at 
[https://cmake.org/download/](https://cmake.org/download/). Make sure to tell 
the installer to add `CMake` to your "PATH".

#### 1.1.2 - Rtools

+ **Option 1**

Use the [`winget`](https://learn.microsoft.com/en-us/windows/package-manager/winget/) 
utility to install `Cmake` from the terminal/PowerShell:

```{powershell eval=FALSE}
winget install RProject.Rtools
```

+ **Option 2**

Use the convenience function provided by the `installr` library: 

```{r eval=FALSE}
if (!require("installr"))
  install.packages("installr")
  
installr::install.Rtools()
```

During the installation process, make sure to tell the installer to add `Rtools` 
to your "PATH".

+ **Option 3**

Download the latest "recommended" version of `Rtools` at 
[https://cran.r-project.org/bin/windows/Rtools/](https://cran.r-project.org/bin/windows/Rtools/). 
Make sure to tell the installer to add `Rtools` to your "PATH".

---

### 1.2 - Mac

Before installing `ROpenCVLite`, you will need to install the latest version of
[`CMake`](https://cmake.org/). 

+ **Option 1**

If you have [`homebrew`](https://brew.sh/) installed, run the following command
in your terminal: 

```bash
brew install cmake
```

+ **Option 2**

If you have [`MacPorts`](https://www.macports.org/) installed, run the following 
command in your terminal: 

```bash
sudo port install cmake
```

+ **Option 3**

Go to [https://cmake.org/download/](https://cmake.org/download/). Scroll down 
until you reach the "Latest Release" section. Download the installer for Mac OS 
X (.dmg file). Once the download is complete, double-click on the installer and 
accept the license agreement. Drag and drop the `Cmake` app into your 
`Applications` folder. 

To make `Cmake` accessible to `R`, open a terminal and type:

`sudo "/Applications/CMake.app/Contents/bin/cmake-gui" --install`

It will ask for your administrator password. 

--- 

### 1.3 - Linux

Before installing `ROpenCVLite`, you will need to install the latest version of
[`CMake`](https://cmake.org/). Hereafter are two options to do this on Linux.

+ **Option 1**

Use your distribution's package management system. For instance on Ubuntu-like
systems, you can install Cmake by typing the following command in your terminal:

```bash
sudo apt install cmake
```

Other distributions will have different package managers. Refer to the 
documentation of your distribution for finding the proper way to install Cmake 
on your system. 

+ **Option 2**

Download `CMake` for Linux at [https://cmake.org/download/](https://cmake.org/download/).
However it is recommended that you install it using your distribution's package
management system.

---

### 1.4 - Mac, Linux and Windows

Before installing `ROpenCVLite`, you will need to install the `devtools` package 
in `R`. 

```{r eval=FALSE}
install.packages("devtools")
```

---

## 2 - Installing `ROpenCVLite`

You can install `ROpenCVLite` from CRAN as follows:

```{r eval=FALSE}
install.packages("ROpenCVLite")
```

Or you can install the latest latest development version (not recommended) as 
follows: 

```{r eval=FALSE}
devtools::install_github("swarm-lab/ROpenCVLite")
```

This will not install `OpenCV` yet. You will be asked whether you want to install
it the first time you load the package. 

```{r eval=FALSE}
library(ROpenCVLite)
```

This step can take some time. Go make yourself a cup of coffee or two while 
`ROpenCVLite` downloads, compiles and installs `OpenCV` on your system. 

In addition, you can also request the (re)installation of `OpenCV` at any time
as follows:

```{r eval=FALSE}
ROpenCVLite::installOpenCV()
```

`ROpenCVLite` will also automatically detect when a new version of OpenCV is 
available and will offer to update it for you. 
