---
author: "Simon Garnier"
title: "1 - Installing ROpenCVLite"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{install}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Installing [`ROpenCVLite`](https://github.com/swarm-lab/ROpenCVLite) should go 
smoothly on all major operating systems (Mac, Windows, "Ubuntu-like" Linux). 
There are a few steps to follow before installing the package itself in order to
install all the necessary external (i.e. non-R) dependencies for 
[`OpenCV`](http://opencv.org/). 

All installation steps are described below. 

---

+ [1 - Before installing ](#before-installing )
  + [1.1 - Windows](#windows)
  + [1.2 - Mac](#mac)
  + [1.3 - Linux](#linux)
  + [1.4 - Mac, Linux and Windows](#mac-linux-and-windows)
+ [2 - Installing `ROpenCVLite`](#installing-ropencvlite)
  
---

## 1 - Before installing 

### 1.1 - Windows

Before installing `ROpenCVLite`, you will need to install the latest versions of
[`CMake`](https://cmake.org/) and [`Rtools`](https://cran.r-project.org/bin/windows/Rtools/). 

Download `CMake` for Windows at [https://cmake.org/download/](https://cmake.org/download/).

Download the latest "recommended" version of `Rtools` at 
[https://cran.r-project.org/bin/windows/Rtools/](https://cran.r-project.org/bin/windows/Rtools/).

In both cases, make sure to tell the installer to add `CMake` and `Rtools` to 
your "PATH".

### 1.2 - Mac

Before installing `ROpenCVLite`, you will need to install the latest version of
[`CMake`](https://cmake.org/). Hereafter are 3 options to do this on Mac.

**Download installer from cmake.org** 

Go to [https://cmake.org/download/](https://cmake.org/download/). Scroll down 
until you reach the "Latest Release" section. Download the installer for Mac OS X
(.dmg file). Once the download is complete, double-click on the installer and 
accept the license agreement. Drag and drop the `Cmake` app into the `Applications`
folder. 

To make `Cmake` accessible to `R`, open a terminal and type:

`sudo "/Applications/CMake.app/Contents/bin/cmake-gui" --install`

It will ask for your administrator password. 

**Homebrew:**

```bash
brew install cmake
```

**MacPorts:**

```bash
sudo port install cmake
```

### 1.3 - Linux

Before installing `ROpenCVLite`, you will need to install the latest version of
[`CMake`](https://cmake.org/). Hereafter are two options to do this on Linux.

**Download installer from cmake.org** 

Download `CMake` for Linux at [https://cmake.org/download/](https://cmake.org/download/).
However it is recommended that you install it using your distribution's package
management system.

**Use you distribution's package management system**

*Ubuntu:*

```bash
sudo apt-get install cmake
```

[TODO: add more Linux install commands]

### 1.4 - Mac, Linux and Windows

Before installing `ROpenCVLite`, you will need to install the `devtools` package 
in `R`. 

```{r eval=FALSE}
install.package("devtools")
```

---

## 2 - Installing `ROpenCVLite`

You can install the latest tested version of `ROpenCVLite` from its Github 
repository as follows:

```{r eval=FALSE}
devtools::install_github("swarm-lab/ROpenCVLite", ref = "v0.3.410")
```

The latest development version (not necessarily tested) can be installed as 
follows: 

```{r eval=FALSE}
devtools::install_github("swarm-lab/ROpenCVLite")
```

This will not install `OpenCV` yet. You will be asked whether you want to install
it the first time you load the package. 

```{r eval=FALSE}
library(ROpenCVLite)
```

This step can take some time. Go make yourself a cup of coffee or two while 
`ROpenCVLite` downloads, compiles and installs `OpenCV` on your system. 

In addition, you can also request the (re)installation of `OpenCV` at any time
as follows:

```{r eval=FALSE}
ROpenCVLite::installOpenCV()
```
